Modal Abstractions for Virtualizing Memory Addresses

### 2023-04-04 programlogic.tex
+

\section{Program Logic for Location Virtualization}
\label{sec:logic}
% The predicate gen_heap_interp.
\newcommand{\gammaPred}{\delta}

\newcommand{\sumwalkabs}[3]{
  \ownGhost\gammaPred{\authfrag{\singletonMap{#1}{(#2, #3)}}}

@@

\begin{figure*}
  \[
  \begin{array}{l}
\begin{array}{l}
  \vaddr\mapsto_{\textsf{t}}\{\textsf{q}\}\vpage \stackrel{def}{=} \lambda \crthree.\\
  \exists_{\entryf \;, \entrytr \;, \entrytw \;,\entryo} \ldotp

  \ulcorner \textsf{aligned } \vaddr \urcorner \ast  \\
  \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\crthree,\entryf,\entrytr,\entrytw,\entryo)\\
   \nfpointsto{\mask\vaddr\ft\entryo}{\mask\vaddr\tw\entryo}\vpage\qfracone\naddr 
\end{array} \\

\textsf{where   }\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr,\crthree,\entryf,\entrytr,\entrytw,\entryo) \stackrel{\triangle}{=} \\
 \qquad\qquad \nfpointsto{\mask\vaddr\ft\crthree}{\mask\vaddr\tw\crthree}\entryf\qfracfotsss\naddr \ast \\ 
 \qquad\qquad  \nfpointsto{\mask\vaddr\ft\entryf}{\mask\vaddr\tw\entryf}\entrytr\qfracfotss\naddr  \ast  \\
  \qquad\qquad \nfpointsto{\mask\vaddr\ft\entrytr}{\mask\vaddr\tw\entrytr}\entrytw\qfracfots\naddr \ast \\
  \qquad\qquad \nfpointsto{\mask\vaddr\ft\entrytw}{\mask\vaddr\tw\entrytw}\entryo\qfracfot\naddr 
   \end{array}
\]
\caption{A Strong Virtual Points-to Relation}
\todo[inline]{These fractions aren't quite right (though I see you added the variance between levels), I can walk you through in our meeting tomorrow.}
  \label{fig:strongvirtualpointsto}
\end{figure*}
\end{remark}
%then we reach to a physical page addr (e.g. \textsf{pa}) 
%\[  \paddr\mapsto_{\textsf{a}}\{1\}\vsome \ast \nfpointsto{\mask\vaddr\ft\entryo}{\mask\vaddr\tw\entryo}\vpage\qfrac\naddr \]

  \subsection{Aliasing/Sharing Physical Pages}
  \label{sec:sharingpages}  
  One might have already observed that the virtual points-to definition shown in Figure \ref{fig:strongvirtualpointsto} is too strong to specify virtual memory operations updating a page table entry. To speak more concretely, the fractional permissions on the page-table-walk assertion ($textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ in Figure \ref{fig:strongvirtualpointsto}) \textit{only} assure the existence of the certain number of mappings -- 512 at each level -- abstracting virtual address translations, and updating any of these mappings would make the other mappings, address translations sharing the page table mappings. Therefore, we need to have logical abstractions ensuring sound updates to the page-table entries in the existence of sharing.

 One possible approach is to abstract away the physical page-table-walk pointsto ($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ in Figure \ref{fig:strongvirtualpointsto}) from the virtual pointsto relation via replacing it with read-only abstract map entries -- a ghost (abstract) map entry per \textsf{L}$_4$\_\textsf{L}$_{1}$ mappings -- which can be used as a resource ($ \sumwalkabs\vaddr\qfrac\paddr$) whose ownership ensures that the virtual address ($\vaddr$) is mapped to a page (referenced by $\paddr$) ,and as a token to access to the physical page-table-walk pointsto (\textsf{L}$_4$\_\textsf{L}$_{1}$) stored in an address-space invariant ($\mathcal{I}$\textsf{ASpace} in Figure \ref{fig:peraspaceinvariant}). We use \textsf{Iris}'s authorative ghost-map logical machinery
\[\mathcal{A}\textsf{bsPTableWalk} \stackrel{\triangle}{=} \textsc{Auth} (\; \mathcal{W}_{64} \;\rightarrow_{\textrm{fin}} \;  (\mathcal{W}_{64} \; \rightarrow_{\textrm{fin}}\;  (\textsc{Frac }, \mathord{+}) \times (\textsc{Agree } \Loc,\mathord{=}) ))\]
to realize these reasoning principles ($\mathcal{A}\textsf{bsPTableWalk}$).
\begin{remark}[Exchanging Page-Table-Walk-Summarization Mappings (Tokens) with Physical Page-Table-Walk Pointstos ]
  $\mathcal{A}\textsf{PTableWalk}$'s fragmental-ownership represents an individual key-value pair (per-virtual-address) which can be handled as a \textit{look-up} resource for an individual virtual address.
  To build more intuition, we can think of the use-case \textit{unmapping-pages} as collecting the fragmental map pairs, e.g. $\sumwalkabs\vaddr\qfrac\paddr$, from collecting the assertions, removing them from the ghost map, and then updating the page tables.
  Critically, physical full ownership ($\ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}}$) of the page tables, as a \textit{singleton resource}, lives in the central invariant keeping physical-page-table pointstos per-virtual address translation.
  With this mechanism we can separate the virtual-to-physical translation from the physical memory involved, and consequently
  \begin{itemize}
  \item virtual pointsto assertions, whose definition is shown in Figure \ref{fig:virtualpointstosharing}, can refer to their physical memory assertions representing their page-address translation in the physical-page-tables ($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$) which reside in the address-space invariant ($\mathcal{I}$\textsf{ASpace} in Figure \ref{fig:peraspaceinvariant}) using individual key-value pair, i.e. instances of ($\sumwalkabs\vaddr\qfrac\paddr$)
  \item unlike the case in the strong virtual-pointsto relation in Figure \ref{fig:strongvirtualpointsto}, physical-page-table-walk pointsto assertions ($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$) for all virtual addresses are moved into the address-space invariant ($\mathcal{I}$\textsf{ASpace} in Figure \ref{fig:peraspaceinvariant}), and resides with authorative ownership of $\mathcal{A}\textsf{PTableWalk}$ ($\ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}}$) which respects all entries of the abstract virtual-address translation mappings ($\sumwalkabs\vaddr\qfrac\paddr$).
    \end{itemize}
\end{remark}
\begin{figure*}
\[
\begin{array}{l}
    \vaddr\mapsto_{\textsf{v}}\{\textsf{q}\}\vpage \stackrel{\triangle}{=} 
  \exists \entryo \ldotp 
  \sumwalkabs\vaddr\qfrac\paddr \ast 
   \nfpointsto{\mask\vaddr\ft\entryo}{\mask\vaddr\tw\entryo}\vpage\qfracone\naddr
\end{array}
\]
\caption{Virtual-Pointsto for Sharing Pages}
  \label{fig:virtualpointstosharing}
\end{figure*}
%\subsection{Address-Space Abstraction}
%\label{sec:aspacemodal}

>>>>

\section{Program Logic for Location Virtualization}
\label{sec:logic}
% The predicate gen_heap_interp.
\newcommand{\gammaPred}{\delta}
\newcommand{\gammaPreds}{\delta\textsf{s}}

\newcommand{\sumwalkabs}[3]{
  \ownGhost\gammaPred{\authfrag{\singletonMap{#1}{(#2, #3)}}}
}

\newcommand{\sumapaces}[2]{
  \ownGhost\gammaPreds{\authfrag{\singletonMap{#1}{#2}}}

@@

\begin{figure*}
  \[
  \begin{array}{l}
\begin{array}{l}
  \vaddr\mapsto_{\textsf{t}}\{\textsf{q}\}\vpage\; : \mathsf{vProp} \; \Sigma \;  \stackrel{def}{:=}  \lambda \crthree. \;\; \\
  \exists_{\entryf \;, \entrytr \;, \entrytw \;,\entryo} \ldotp

  \ulcorner \textsf{aligned } \vaddr \urcorner \ast  \\
 \quad \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\crthree,\entryf,\entrytr,\entrytw,\entryo) \ast \nfpointsto{\mask\vaddr\ft\entryo}{\mask\vaddr\tw\entryo}\vpage\qfracone\naddr 
\end{array} \\
\\
\textsf{where   }\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr,\crthree,\entryf,\entrytr,\entrytw,\entryo) \stackrel{\triangle}{=} \\
 \qquad\qquad \nfpointsto{\mask\vaddr\ft\crthree}{\mask\vaddr\tw\crthree}\entryf\qfracfotsss\naddr \ast \\ 
 \qquad\qquad  \nfpointsto{\mask\vaddr\ft\entryf}{\mask\vaddr\tw\entryf}\entrytr\qfracfotss\naddr  \ast  \\
  \qquad\qquad \nfpointsto{\mask\vaddr\ft\entrytr}{\mask\vaddr\tw\entrytr}\entrytw\qfracfots\naddr \ast \\
  \qquad\qquad \nfpointsto{\mask\vaddr\ft\entrytw}{\mask\vaddr\tw\entrytw}\entryo\qfracfot\naddr 
   \end{array}
\]
\caption{A Strong Virtual Points-to Relation}
\todo[inline]{These fractions aren't quite right (though I see you added the variance between levels), I can walk you through in our meeting tomorrow.}
  \label{fig:strongvirtualpointsto}
\end{figure*}
\end{remark}
%then we reach to a physical page addr (e.g. \textsf{pa}) 
%\[  \paddr\mapsto_{\textsf{a}}\{1\}\vsome \ast \nfpointsto{\mask\vaddr\ft\entryo}{\mask\vaddr\tw\entryo}\vpage\qfrac\naddr \]

  \subsection{Aliasing/Sharing Physical Pages}
  \label{sec:sharingpages}  
  One might have already observed that the virtual points-to definition shown in Figure \ref{fig:strongvirtualpointsto} is too strong to specify virtual memory operations updating a page table entry. To speak more concretely, the fractional permissions on the page-table-walk assertion ($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ in Figure \ref{fig:strongvirtualpointsto}) \textit{only} assure the existence of the certain number of mappings -- 512 at each level -- abstracting virtual address translations, and updating any of these mappings would make the other mappings, address translations sharing the page table mappings. Therefore, we need to consturct logical abstractions ensuring sound updates to the page-table entries in the existence of sharing.

 One possible approach is to abstract away the physical page-table-walk pointsto ($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ in Figure \ref{fig:strongvirtualpointsto}) from the virtual pointsto relation via replacing it with read-only abstract map entries -- a ghost (abstract) map entry per \textsf{L}$_4$\_\textsf{L}$_{1}$ mappings -- which can be used as a resource ($ \sumwalkabs\vaddr\qfrac\paddr$) whose ownership ensures that the virtual address ($\vaddr$) is mapped to a page (referenced by $\paddr$) ,and as a token to access to the physical page-table-walk pointsto (\textsf{L}$_4$\_\textsf{L}$_{1}$) stored in an address-space invariant ($\mathcal{I}$\textsf{ASpace} in Figure \ref{fig:peraspaceinvariant}). We use \textsf{Iris}'s authorative ghost-map logical machinery
\[\mathcal{A}\textsf{bsPTableWalk} \stackrel{\triangle}{=} \textsc{Auth} (\; \mathcal{W}_{64} \;\rightarrow_{\textrm{fin}} \;  (\mathcal{W}_{64} \; \rightarrow_{\textrm{fin}}\;  (\textsc{Frac }, \mathord{+}) \times (\textsc{Agree } \Loc,\mathord{=}) ))\]
to realize these reasoning principles ($\mathcal{A}\textsf{bsPTableWalk}$).
\begin{remark}[Exchanging Page-Table-Walk-Summarization Mappings (Tokens) with Physical Page-Table-Walk Pointstos ]
  $\mathcal{A}\textsf{PTableWalk}$'s fragmental-ownership represents an individual key-value pair (per-virtual-address) which can be handled as a \textit{look-up} resource for an individual virtual address.
  To build more intuition, we can think of the use-case \textit{unmapping-pages} as collecting the fragmental map pairs, e.g. $\sumwalkabs\vaddr\qfrac\paddr$, from collecting the assertions, removing them from the ghost map, and then updating the page tables.
  Critically, physical full ownership ($\ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}}$) of the page tables, as a \textit{singleton resource}, lives in the central invariant keeping physical-page-table pointstos per-virtual address translation.
  With this mechanism we can separate the virtual-to-physical translation from the physical memory involved, and consequently
  \begin{itemize}
  \item virtual pointsto assertions, whose definition is shown in Figure \ref{fig:virtualpointstosharing}, can refer to their physical memory assertions representing their page-address translation in the physical-page-tables ($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$) which reside in the address-space invariant ($\mathcal{I}$\textsf{ASpace} in Figure \ref{fig:peraspaceinvariant}) using individual key-value pair, i.e. instances of ($\sumwalkabs\vaddr\qfrac\paddr$)
  \item unlike the case in the strong virtual-pointsto relation in Figure \ref{fig:strongvirtualpointsto}, physical-page-table-walk pointsto assertions ($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$) for all virtual addresses are moved into the address-space invariant ($\mathcal{I}$\textsf{ASpace} in Figure \ref{fig:peraspaceinvariant}), and resides with authorative ownership of $\mathcal{A}\textsf{PTableWalk}$ ($\ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}}$) which respects all entries of the abstract virtual-address translation mappings ($\sumwalkabs\vaddr\qfrac\paddr$).
    \end{itemize}
\end{remark}
\begin{figure*}
\[
\begin{array}{l}
    \vaddr\mapsto_{\textsf{v}}\{\textsf{q}\}\vpage \stackrel{\triangle}{=} 
    \exists \entryo \ldotp
    \sumapaces\crthree\delta \ast 
  \sumwalkabs\vaddr\qfrac\paddr \ast 
   \nfpointsto{\mask\vaddr\ft\entryo}{\mask\vaddr\tw\entryo}\vpage\qfracone\naddr
\end{array}
\]
\caption{Virtual-Pointsto for Sharing Pages}
  \label{fig:virtualpointstosharing}
\end{figure*}
%\subsection{Address-Space Abstraction}
%\label{sec:aspacemodal}
